---
  title: "Part - II"
author: "Surya vamsi Patiballa"
date: "`r Sys.Date()`"
output:
  html_document:
  code_folding: hide
number_sections: false
toc: yes
toc_depth: 3
toc_float: yes
pdf_document:
  toc: yes
toc_depth: '3'
---
  
```{r init, include=F}
hospital_data <- read.csv("C:\Users\HP\Desktop\Class files\Data science\Project - 1\dataset.csv")
head(hospital_data)

```

#How does the presence of diabetes mellitus correlate with age? Is diabetes a risk factor for hospital
#death in older patients, and does age significantly impact the chance of developing diabetes?

```{r setup, include=FALSE}
library(dplyr)
library(stringr)

# Check the structure of the dataset
str(hospital_data)
correlation_age_diabetes <- cor(hospital_data$age, hospital_data$diabetes_mellitus)
cat("Correlation between Age and Diabetes Mellitus:", correlation_age_diabetes, "\n")
hospital_data <- hospital_data %>%
  mutate(age_group = ifelse(age >= 65, "Older", "Younger"))
cross_table <- table(hospital_data$diabetes_mellitus, hospital_data$age_group, hospital_data$hospital_death)
print(cross_table)

correlation_age_diabetes <- cor(hospital_data$age, hospital_data$diabetes_mellitus)
cat("Correlation between Age and Diabetes Mellitus:", correlation_age_diabetes, "\n")
hospital_data <- hospital_data[!is.na(hospital_data$age) & is.numeric(hospital_data$age), ]
logistic_model <- glm(hospital_death ~ age + diabetes_mellitus, data = hospital_data, family = "binomial")

# Create a sequence of ages for plotting purposes
age_seq <- seq(min(hospital_data$age), max(hospital_data$age), length.out = 100)

# Create a data frame for predictions
new_data <- expand.grid(age = age_seq, diabetes_mellitus = c(0, 1))

# Predict probabilities
new_data$predicted_prob <- predict(logistic_model, newdata = new_data, type = "response")

# Plot logistic regression curve with scatter plot
library(ggplot2)

ggplot(hospital_data, aes(x = age, y = hospital_death)) +
  geom_point(aes(color = factor(diabetes_mellitus)), alpha = 0.6) +
  geom_line(data = new_data, aes(y = predicted_prob), color = "blue", size = 1) +
  scale_color_discrete(name = "Diabetes Mellitus") +
  labs(x = "Age", y = "Hospital Death Probability") +
  ggtitle("Logistic Regression: Probability of Hospital Death by Age and Diabetes Mellitus") +
  theme_minimal()

# Fit logistic regression model
logistic_model <- glm(hospital_death ~ age + diabetes_mellitus, data = hospital_data, family = "binomial")
summary(logistic_model)
```

#Are there any particular illnesses or risk factors (such as leukemia, cirrhosis, or diabetes mellitus) that
#are closely linked to hospital deaths?
  
```{r}

selected_columns <- hospital_data %>%
  select(hospital_death, leukemia, cirrhosis, diabetes_mellitus)
model_leukemia <- glm(hospital_death ~ leukemia, data = selected_columns, family = "binomial")
model_cirrhosis <- glm(hospital_death ~ cirrhosis, data = selected_columns, family = "binomial")
model_diabetes <- glm(hospital_death ~ diabetes_mellitus, data = selected_columns, family = "binomial")

summary(model_leukemia)
summary(model_cirrhosis)
summary(model_diabetes)

library(ggplot2)
library(dplyr)

# Select relevant columns for analysis (e.g., leukemia, cirrhosis, diabetes_mellitus, hospital_death)
selected_data <- hospital_data %>%
  select(leukemia, cirrhosis, diabetes_mellitus, hospital_death)  # Update column names as needed

# Calculate hospital death rates for each condition
death_rates <- selected_data %>%
  group_by(leukemia, cirrhosis, diabetes_mellitus) %>%
  summarize(death_rate = mean(hospital_death == 1))

# Reshape the data for plotting
death_rates_long <- death_rates %>%
  pivot_longer(cols = c(leukemia, cirrhosis, diabetes_mellitus),
               names_to = "Condition",
               values_to = "Death_Rate")

# Create a bar plot showing death rates for different conditions
ggplot(death_rates_long, aes(x = Condition, y = Death_Rate, fill = Condition)) +
  geom_bar(stat = "identity") +
  labs(x = "Condition", y = "Hospital Death Rate", title = "Hospital Death Rates by Condition") +
  theme_minimal()
```
#Which underlying medical disorders account for the majority of hospital deaths among patients over the
#age of 70? Are there any trends in the kinds of illnesses that cause death in this particular age range?
  ```{r}

library(ggplot2)
library(dplyr)

# Filter data for patients over the age of 70 who experienced hospital death
older_patients_death <- hospital_data %>%
  filter(age > 70, hospital_death == 1)

# Select relevant columns related to medical disorders or conditions (modify as needed)
selected_columns <- older_patients_death %>%
  select(diabetes_mellitus, cirrhosis, leukemia, heart_rate_apache,glucose_apache)  # Add other relevant columns

# Calculate the count of each medical condition
condition_counts <- selected_columns %>%
  summarise_all(sum) %>%
  pivot_longer(cols = everything(), names_to = "Condition", values_to = "Count")

# Create a bar plot to visualize medical conditions among deceased patients over 70
ggplot(condition_counts, aes(x = reorder(Condition, -Count), y = Count, fill = Condition)) +
  geom_bar(stat = "identity") +
  labs(x = "Medical Condition", y = "Count", title = "Medical Conditions among Deceased Patients over 70") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
#What is the mortality rate in the dataset between patients under 65 and those above 65? Are the
#causes of death for these age groups significantly different from one another?
  
```{r}

library(dplyr)

# Calculate mortality rates for patients under 65 and above 65
hospital_data$age_group <- ifelse(hospital_data$age < 65, "Under 65", "Above 65")

# Check the structure of 'cause_of_death_column' to ensure it has the same length as other columns
str(hospital_data$cause_of_death_column)



# Perform chi-squared test for independence
chi_square_test <- chisq.test(contingency_table)
print(chi_square_test)

ggplot(mortality_rates, aes(x = age_group, y = mortality_rate, fill = age_group)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Age Group", y = "Mortality Rate", title = "Mortality Rates by Age Group") +
  theme_minimal()


```
#Which ICU types are most prevalent in the dataset, and does the type of ICU affect the risk of hospital
#death?
  
```{r}

library(dplyr)

# Calculate frequency of each ICU type
icu_type_counts <- hospital_data %>%
  group_by(icu_type) %>%
  summarise(count = n()) %>%
  arrange(desc(count))

# Display ICU type counts
print(icu_type_counts)

death_rates_by_icu <- hospital_data %>%
  group_by(icu_type) %>%
  summarise(death_rate = mean(hospital_death == 1, na.rm = TRUE))

# Display hospital death rates by ICU type
print(death_rates_by_icu)

library(ggplot2)

ggplot(death_rates_by_icu, aes(x = icu_type, y = death_rate)) +
  geom_bar(stat = "identity", fill = "skyblue", color = "black") +
  labs(x = "ICU Type", y = "Hospital Death Rate", title = "Hospital Death Rates by ICU Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
#What differences exist between patients who survive and those who do not in terms of the distribution
#of "BMI" (body mass index)? Does a person's BMI have any bearing on hospital mortality?

```{r}

survived_bmi <- hospital_data$hospital_death == 0  # Assuming 0 represents survived patients
not_survived_bmi <- hospital_data$hospital_death == 1  # Assuming 1 represents not survived patients

summary(hospital_data$BMI[survived_bmi])  # Summary stats for survived patients' BMI
summary(hospital_data$BMI[not_survived_bmi])  # Summary stats for not survived patients' BMI

boxplot(bmi ~ hospital_death, data = hospital_data, xlab = "Hospital Mortality", ylab = "BMI")

t_test_result <- t.test(hospital_data$bmi[survived_bmi], hospital_data$BMI[not_survived_bmi])
print(t_test_result)
```
#Are there significant differences in the distribution of vital signs (heart rate, glucose level) between different ethnic groups in the dataset?
```{r}

library(dplyr)
library(ggplot2)

# Subset the dataset with relevant columns
subset_data <- hospital_data %>%
  select(ethnicity, heart_rate_apache, glucose_apache)

# Boxplot visualization of heart rate by ethnicity
ggplot(subset_data, aes(x = ethnicity, y = heart_rate_apache)) +
  geom_boxplot() +
  labs(x = "Ethnicity", y = "Heart Rate") +
  ggtitle("Distribution of Heart Rate by Ethnicity")

# Boxplot visualization of glucose level by ethnicity
ggplot(subset_data, aes(x = ethnicity, y = glucose_apache)) +
  geom_boxplot() +
  labs(x = "Ethnicity", y = "Glucose Level") +
  ggtitle("Distribution of Glucose Level by Ethnicity")

# Perform statistical tests (e.g., ANOVA or Kruskal-Wallis) to compare vital signs among ethnic groups
# For heart rate:
heart_rate_anova <- aov(heart_rate_apache ~ ethnicity, data = subset_data)
print(summary(heart_rate_anova))  # ANOVA test

# For glucose level:
glucose_kruskal <- kruskal.test(glucose_apache ~ ethnicity, data = subset_data)
print(glucose_kruskal)
```
#How does the prevalence of specific pre-existing conditions (diabetes, cirrhosis, leukemia) vary among different age groups, and is there a correlation between age and the occurrence of these conditions?

```{r}

library(dplyr)
library(ggplot2)
library(tidyr) 

# Select relevant columns for analysis
subset_data <- hospital_data %>%
  select(age, diabetes_mellitus, cirrhosis, leukemia)

# Calculate prevalence of each condition by age group
age_groups <- cut(subset_data$age, breaks = c(0, 18, 40, 60, 80, 100), labels = c("0-18", "19-40", "41-60", "61-80", "81-100"))
subset_data <- subset_data %>%
  mutate(age_group = age_groups)

# Calculate the proportion of each condition within each age group
condition_prevalence <- subset_data %>%
  group_by(age_group) %>%
  summarize(
    diabetes_prevalence = mean(diabetes_mellitus),
    cirrhosis_prevalence = mean(cirrhosis),
    leukemia_prevalence = mean(leukemia)
  )

# Plotting condition prevalence by age group
condition_plot <- gather(condition_prevalence, condition, prevalence, -age_group) %>%
  ggplot(aes(x = age_group, y = prevalence, fill = condition)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Age Group", y = "Prevalence", fill = "Condition") +
  ggtitle("Prevalence of Conditions by Age Group") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(condition_plot)

# Correlation analysis between age and conditions
correlation_matrix <- cor(subset_data[, c("age", "diabetes_mellitus", "cirrhosis", "leukemia")])
print(correlation_matrix)
```
#Can we analyze the trends in hospital mortality rates over time? Is there a noticeable change or improvement in mortality rates in recent years?

```{r}

library(dplyr)
library(ggplot2)

# Assuming there's a timestamp column indicating admission or discharge time
# Convert timestamp to a Date format if needed
hospital_data$timestamp <- seq(as.Date("2020-01-01"), by = "1 day", length.out = nrow(hospital_data))  # Replace 'timestamp' with the actual column name

# Calculate hospital mortality rate per year
mortality_rate <- hospital_data %>%
  mutate(year = lubridate::year(timestamp)) %>%  # Extract year from the timestamp
  group_by(year) %>%
  summarize(
    mortality_rate = mean(hospital_death)  # Calculate mean hospital mortality rate per year
  )

# Plotting the trend of hospital mortality rates over time
mortality_plot <- ggplot(mortality_rate, aes(x = year, y = mortality_rate)) +
  geom_line() +
  labs(x = "Year", y = "Mortality Rate", title = "Trends in Hospital Mortality Rates Over Time")

print(mortality_plot)
```
#Are there specific combinations of pre-existing conditions or a particular set of vital signs that indicate a higher likelihood of hospital mortality?

```{r}

library(dplyr)

library(caret)

colnames(hospital_data)
head(hospital_data)
# Select relevant columns for analysis (pre-existing conditions and vital signs)
subset_data <- hospital_data %>%
  select(age, gender, diabetes_mellitus, cirrhosis, heart_rate_apache, glucose_apache, hospital_death) # Add more vital signs or conditions as needed

# Handle missing values if any
subset_data <- na.omit(subset_data)

# Split data into training and testing sets
set.seed(123)
trainIndex <- createDataPartition(subset_data$hospital_death, p = 0.7, list = FALSE)
train_data <- subset_data[trainIndex, ]
test_data <- subset_data[-trainIndex, ]

# Fit logistic regression model
logistic_model <- glm(hospital_death ~ ., data = train_data, family = "binomial")

# Summary of the model
summary(logistic_model)

# Predict hospital mortality on test data
predicted <- predict(logistic_model, newdata = test_data, type = "response")

# Evaluate model performance (e.g., using AUC-ROC)
library(pROC)
auc_score <- roc(test_data$hospital_death, predicted)
print(auc_score)
```